Hypertension and microbiome

### 2019-06-14 articletwo.Rmd
Cleaner salt data frame creation

```{r libraries, cache = FALSE}
library(dplyr)
library(tibble)
library(phyloseq)
library(nortest)
library(microbiome)
library(knitr)
library(tidyr)
library(vegan)
library(reshape)
library(parallel)
library(officer)
library(flextable)
library(xtable)
library(rvg)
library(tableone)
library(scales)
library(ggplot2)
library(gridExtra)
library(png)
library(pander)
library(ggpubr)
library(broom)
library(ggfortify)
library(RColorBrewer)
library(gvlma)
library(purrr)
library(pwr)
library(gtable)
```

@@

```{r maaslin runs, results = 'hide'} 
if (!file.exists("rds/maaslin.rds")) {
    pseq.genus.core <- core(pseq.genus, detection = 0.001, prevalence = 0.01)
    pseq.genus.core.viruses.names <- taxa(pseq.genus.core) %>% mygrep(word = "virus$", complement = TRUE)
    pseq.genus.core.pruned <- prune_taxa(pseq.genus.core.viruses.names, pseq.genus.core)

    maaslin <- maaslinwrapper(pseq = pseq.genus.core.pruned,
                                    looped = var.BP,
                                    forced = var.CL,
                              taxa = taxa(pseq.genus.core.pruned))

@@

pcoa.df %>% pull(Axis.1) %>% summary

@@

@@

```{r lacto data frame}
pseq.genus.salt <- pseq.genus %>%
    prune_samples(rownames(subset(meta(.), !is.na(NA.))), .) %>%
    prune_taxa(c("g_Lactobacillus", "g_Lactococcus"), .)

abund.genus.salt <- pseq.genus.salt %>%
    abundances(., transform = "clr") %>%
    t %>%
    as.data.frame %>%
    tibble::rownames_to_column("id") %>%
    select(-g_Lactococcus)

df.genus.salt <- meta(pseq.genus.salt) %>%
    tibble::rownames_to_column("id") %>%
    full_join(., abund.genus.salt, by = "id") %>%
    dplyr::mutate(dUNA = myscale(NA.),
                  salt_groups = relevel(factor(dplyr::ntile(dUNA, 3)), ref='1'),
                  lacto_groups = relevel(factor(dplyr::ntile(g_Lactobacillus, 3)), ref='1'))
```

@@

Bartlett's test for equality of variance gives p = `r
bartlett.test(g_Lactobacillus ~ salt_groups, data =
df.genus.salt)$p.value`.

One-way test assuming normal distributions gives for equal mean p = `r
oneway.test(g_Lactobacillus ~ salt_groups, data =
df.genus.salt, var.equal = TRUE)$p.value`.

@@

>>>>

```{r libraries, cache = FALSE}
library(dplyr)
library(tibble)
library(phyloseq)
library(nortest)
library(microbiome)
library(knitr)
library(tidyr)
library(vegan)
library(reshape)
library(parallel)
library(officer)
library(flextable)
library(xtable)
library(rvg)
library(tableone)
library(scales)
library(ggplot2)
library(gridExtra)
library(png)
library(pander)
library(ggpubr)
library(broom)
library(ggfortify)
library(RColorBrewer)
library(gvlma)
library(purrr)
library(pwr)
library(gtable)
library(car)
```

@@

```{r maaslin runs, results = 'hide'} 
 pseq.genus.core <- core(pseq.genus, detection = 0.001, prevalence = 0.01)
 pseq.genus.core.viruses.names <- taxa(pseq.genus.core) %>% mygrep(word = "virus$", complement = TRUE)
 pseq.genus.core.pruned <- prune_taxa(pseq.genus.core.viruses.names, pseq.genus.core)

if (!file.exists("rds/maaslin.rds")) {
     maaslin <- maaslinwrapper(pseq = pseq.genus.core.pruned,
                                    looped = var.BP,
                                    forced = var.CL,
                              taxa = taxa(pseq.genus.core.pruned))

@@

@@

```{r maaslin means}
pseq.genus.core.pruned.table <- pseq.genus.core.pruned %>%
    prune_taxa(c("g_Adlercreutzia", "g_Anaerotruncus", "g_Sutterella"), .) %>%
    transform(transform = "compositional") %>%
    psmelt %>%
    mutate(map_groups = relevel(factor(dplyr::ntile(MAP, 3)), ref='1'),
           systm_groups = relevel(factor(dplyr::ntile(SYSTM, 3)), ref='1'),
           Abundance = myscale(Abundance))

table_bp_abund <- function(pseq, group) {
    pseq %>%
        group_by(.dots = c(group, "OTU")) %>%
        summarize(mean = mean(Abundance), sd = sd(Abundance)) %>%
        mutate(beta_sd = sprintf("%.2f ± %.2f", mean, sd)) %>%
        select(OTU, beta_sd) %>%
        spread(OTU, beta_sd) %>%
        ungroup %>%
        dplyr::rename("grp" = group) %>%
        mutate("grp" = paste0(group, "_", grp))
}

rbind(table_bp_abund(pseq.genus.core.pruned.table, "map_groups"),
             table_bp_abund(pseq.genus.core.pruned.table, "systm_groups")) %>%
    tibble::column_to_rownames("grp") %>%
    t %>%
    as.data.frame %>%
    kable

@@

```{r lacto data frame}
df.genus.salt <- pseq.genus %>%
    microbiome::transform(transform = "clr") %>%
    prune_taxa(c("g_Lactobacillus"), .) %>%
    psmelt %>%
    dplyr::filter(!is.na(NA.)) %>%
    spread(OTU, Abundance) %>%
    dplyr::mutate(dUNA = myscale(NA.),
                  salt_groups = relevel(factor(dplyr::ntile(dUNA, 3)), ref='1'),
                  lacto_groups = relevel(factor(dplyr::ntile(g_Lactobacillus, 3)), ref='1'))
```

Lactobacillus prevalence

```{r lactobacillus prevalence}
pseq.genus %>%
    microbiome::transform(transform = "compositional") %>%
    prune_taxa(c("g_Lactobacillus"), .) %>%
    psmelt %>%
    spread(OTU, Abundance) %>%
    dplyr::mutate(lacto.prevalent = ifelse(g_Lactobacillus > 0.001, TRUE, FALSE),
                  duna.present = ifelse(is.na(NA.), 0, 1)) %>%
    group_by(duna.present) %>%
    summarize(lacto_detected = sum(lacto.prevalent), n = n(), ratio = lacto_detected/n)
```

@@

@@

Levene's test for equality of variance

```{r salt levene}
car::leveneTest(g_Lactobacillus ~ salt_groups, data = df.genus.salt) %>% kable
```
